Practical Optional Types for Clojure

### 2015-07-10 metatheory.tex
modified some of proof
update metatheory
edits to 4

We prove type soundness follow using the same technique as~\citet{TF10}. 
We also include errors and a \wrong{} value and prove
well-typed programs do not go wrong.

Rather than modeling Java's dynamic semantics, we instead
make our assumptions about Java explicit. We concede that
method and constructor calls may diverge or error, but we assume they can
never go wrong. (Assumptions for other operations are given in the supplemental
material).

@@

For the purposes of our soundness proof, we require that all values
are \emph{consistent}.
Consistency
ensures that occurrence typing does not refer to variables
hidden inside a closure.

{\consistentwithonlydef{main}}

@@

  {\soundnesslemmahypothesis}
  \begin{proof}
    By induction on the derivation of the typing judgment. 
    (Full proof given as lemma~\ref{appendix:lemma:soundness}).
  \end{proof}
\end{lemma}

We can now state our soundness theorem.

>>>>

We prove type soundness following \citet{TF10}.  Our model is extended
to include errors and a \wrong{} value, and we prove well-typed
programs do not go wrong; this is therefore a stronger theorem than
proved by \citet{TF10}.

Rather than modeling Java's dynamic semantics, a task of daunting
complexity, we instead make our assumptions about Java explicit. We
concede that method and constructor calls may diverge or error, but
assume they can never go wrong. (Assumptions for other operations are
given in the supplemental material).

@@

For the purposes of our soundness proof, we require that all values
are \emph{consistent}.  Consistency (defined in the supplemental
material) states that the types of closures are well-scoped---they do
not claim propositions about variables hidden in their closures.

%{\consistentwithonlydef{main}}

@@

  {\soundnesslemmahypothesis}
  \begin{proof}
    By induction on the derivation of \opsem {\openv{}} {\e{}} {\a{}}.
    (Full proof given as lemma~\ref{appendix:lemma:soundness}).
  \end{proof}
\end{lemma}

We can now state our soundness theorems.
